---
title: "p8105_hw2_gl2761"
author: "Gonghao Liu"
date: "10/4/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```

### Problem 1
```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

These code chunk first import the dataset, then convert the 4 route variables from num to char, `janitor::clean` was use to get a clean column name. Then use `select` function to select the column we want. Finally, update values in entry column from 'yes' and 'no' to 'TRUE' and 'FALSE'. 

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

There are 465 distinct stations here.

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 84 stations are ADA compliant.

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

There are 37.70492% of station entrance / exists without vending allow entrance. 

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 60 distinct stations serve the A train and 17 stations are ADA compliant.

### Problem 2

Read and clean the Mr. Trash Wheel sheet.

```{r trash_wheel, message = FALSE}
trash_wheel = read_excel("data/Trash Wheel Collection Data.xlsx", 
                         sheet = "Mr. Trash Wheel", 
                         range = "A2:N547")%>%
  janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls = as.integer(round(sports_balls))) %>%
  # New column to track sheet
  mutate(sheet = "Mr. Trash Wheel")
```

Read and clean the Porfessor Trash Wheel sheet.

```{r pro_trash_wheel, message = FALSE}
pro_trash_wheel = read_excel("data/Trash Wheel Collection Data.xlsx", 
                         sheet = "Professor Trash Wheel", 
                         range = "A2:M96")%>%
  janitor::clean_names()%>%
  drop_na(dumpster)%>%
  # New column to track sheet
  mutate(sheet = "Professor Trash Wheel") %>%
  # Create a new column and convert the variable to combine the datasets
  mutate(sports_balls = 0) %>%
  mutate(year = as.character(year))
```

Combine the pro_trash_wheel and trash_wheel as one dataset.

```{r combined_trash_wheel, message = FALSE}
combined_trash_wheel = 
  bind_rows(trash_wheel, pro_trash_wheel)
```

Write a paragraph about data:

In Trash Wheel data, it has key variables: `r names(trash_wheel)`, with number of `r ncol(trash_wheel)` and has `r nrow(trash_wheel)` observations.

In Porfessor Trash Wheel data, it has key variables: `r names(pro_trash_wheel)`, with number of `r ncol(pro_trash_wheel)` and has `r nrow(pro_trash_wheel)` observations.

In Combined Trash Wheel data, it has key variables: `r names(combined_trash_wheel)`, with number of `r ncol(combined_trash_wheel)` and has `r nrow(combined_trash_wheel)` observations.

The total weight of trash collected by Professor Trash Wheel is `r sum(pull(pro_trash_wheel, weight_tons))`.

The the total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(pull(filter(trash_wheel, year == 2020), sports_balls))`.

### Problem 3

Read and clean data in pols_month.

```{r pols_month, message = FALSE}
pols_month = read_csv("./fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names()%>%
  separate(mon, into = c("year", "month", "day"))%>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.name[month],
    president = ifelse(prez_dem == "1", "dem", "gop")
    )%>%
  select(- day, -prez_dem, -prez_gop)%>%
relocate(year, month, president)
```

Read and clean the data snp.csv as well, arrange according to year and month, and organize columns.

```{r snp, message = FALSE}
snp = read_csv("./fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names()%>%
  mutate(date = lubridate::mdy(date))%>%
  separate(date, into = c("year", "month", "day"))%>%
  mutate(
   month = as.integer(month),
   month = month.name[month],
   year = as.integer(year),
   year = ifelse(year>2021, year - 100, year)
  )%>%
  select(-day)%>%
  arrange(year, month)%>%
  relocate(year,month)
```

Read and clean the unemployment data.

```{r unemployment_tidy, message = FALSE}
unemployment_data = read_csv("./fivethirtyeight_datasets/unemployment.csv")%>%
  janitor::clean_names()
colnames(unemployment_data) = c("year","January","February","March","April","May","June","July","August","September","October","November","December")

unemployment_tidy = 
  unemployment_data %>%
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemp_percent"
  )%>%
  drop_na()%>%
  relocate(year, month)
```

Join the datasets.

```{r join, message = FALSE}
pols_snp = left_join(pols_month, snp, by = c("year", "month"))

pols_snp_unemp = left_join(pols_snp, unemployment_tidy, by = c("year", "month"))
```

A short paragraph about these data sets.
In our pols_month data set, it shows the number of national politicians who are democratic or republican at any given time. It has `r ncol(pols_month)` variables which are `r names(pols_month)`, and has `r nrow(pols_month)` observations. The time period is from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))`.

In snp data set, it shows Standard & Poor’s stock market index (S&P). It has `r ncol(snp)` variables including `r names(snp)`, and has `r nrow(snp)` observations. The time period is from `r min(pull(snp, year))` to `r max(pull(snp, year))`.

In unemployment data set, it has `r ncol(unemployment_tidy)` variables with the name of `r names(unemployment_tidy)`, and it has `r nrow(unemployment_tidy)` observations. Time range is between `r min(pull(unemployment_tidy, year))` to `r max(pull(unemployment_tidy, year))`.

In the first combined data set, it combined pols_month data and snp data, with dimension `r nrow(pols_snp)` and `r ncol(pols_snp)`. Variables include `r names(pols_snp)`. Time period is from `r min(pull(pols_snp, year))` to `r max(pull(pols_snp, year))`.

The last dataset contains pols_month data, snp data and unemployment data, with `r nrow(pols_snp_unemp)` observations and `r ncol(pols_snp_unemp)` variables. Variables include `r names(pols_snp_unemp)`. Time period is from `r min(pull(pols_snp_unemp, year))` to `r max(pull(pols_snp_unemp, year))`.
